---
title: "gsea fr"
author: "Aditi Goyal"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning  = FALSE)
knitr::opts_chunk$set(message  = FALSE)
```

```{r}
library(fgsea)
library(tidyverse)
library(dplyr)
library(tibble)
library(GO.db)
library(KEGGREST)
```

```
library(DESeq2)
library(EnhancedVolcano)
library(readxl)
library(umap)
library(pheatmap)
library(dendextend)
library(colorspace)
library(clusterProfiler)
library(ggupset)
library(enrichplot)
library(eulerr)
library(pathview)
```

```{r}
rm(list=ls())
setwd('~/Desktop/KAISHA')
```

# Setup
```{r}
full_map=read_csv('mappings/full_map.csv')
G2_G0=read.csv('G2_G0.csv')
G10_G0=read.csv('G10_G0.csv')
G17_G0=read.csv('G17_G0.csv')
H_A=read.csv('heatshock_deseq.csv')
```

```{r}
GO <- as.list(GOTERM)
#go_gmt=gmtPathways('go_uniprot.gmt') #taxon id long one
go_gmt=gmtPathways('GSEA/go_1282.gmt')
```

```{r}
pathways.list <- keggList("pathway", "sep")
pathway.codes <- sub("path:", "", names(pathways.list)) 

kegg_gmt=gmtPathways('kegg.gmt')
```
# Functions



```{r}
deseq_format = function(df){
  colnames(df)[1]='X'
  df=df%>%dplyr::rename('gene_name'='X')
  
  #remove _C suffix
  df$gene_name=gsub(pattern = '_C', replacement = '', df$gene_name)
  
  #remove _P suffix
  df$gene_name=gsub(pattern = '_P', replacement = '', df$gene_name)
  
  if ('lepBP' %in% df$gene_name){
    names=df$gene_name
    names[names=='lepBP']='lepB'
    df$gene_name=names
  }
  
  merged=inner_join(df, full_map, by='gene_name')
  return(merged)
}
```

```{r}
counts_to_fgsea_GO=function(df, p=0.05){
  df_formatted=deseq_format(df)
  
  ranks=df_formatted$log2FoldChange
  names(ranks)=df_formatted$uniprot_quickGO
  ranks=sort(ranks, decreasing = TRUE)
  
  set.seed(12345)
  fgsea_res=fgsea(go_gmt, ranks, minSize = 15, maxSize = 500)
  
  fgsea_res=fgsea_res%>%mutate(overlap = lengths(leadingEdge))%>%mutate(gene_ratio=overlap/size)
  
  sig_fgsea=fgsea_res%>%filter(padj<=p) %>% relocate(padj, .after = pathway)
  
  sig_fgsea=sig_fgsea%>%dplyr::select(pathway, padj, NES, gene_ratio)
  
  fgsea_up=sig_fgsea%>%filter(NES>0)
  fgsea_down=sig_fgsea%>%filter(NES<0)
  
  return(list(fgsea_up, fgsea_down))
}
```


```{r}
counts_to_fgsea_KEGG=function(df, p=0.05){
  df_formatted=deseq_format(df)
  
  ranks=df_formatted$log2FoldChange
  names(ranks)=df_formatted$sseqid
  ranks=sort(ranks, decreasing = TRUE)
  
  set.seed(12345)
  fgsea_res=fgsea(kegg_gmt, ranks, minSize = 15, maxSize = 500)
  
  fgsea_res=fgsea_res%>%mutate(overlap = lengths(leadingEdge))%>%mutate(gene_ratio=overlap/size)
  
  sig_fgsea=fgsea_res%>%filter(padj<=p) %>% relocate(padj, .after = pathway)
  
  sig_fgsea=sig_fgsea%>%dplyr::select(pathway, padj, NES, gene_ratio)
  
  fgsea_up=sig_fgsea%>%filter(NES>0)
  fgsea_down=sig_fgsea%>%filter(NES<0)
  
  return(list(fgsea_up, fgsea_down))
}
```

```{r}
format_gsea_results=function(df, p=0.05){
  
  df$gene_ratio=sub(',.*', '',df$`LEADING EDGE` )

  df$gene_ratio=sub('tags=', '', df$gene_ratio)

  df$gene_ratio=sub('%', '', df$gene_ratio)

df$gene_ratio=as.numeric(df$gene_ratio)

df$gene_ratio=df$gene_ratio/100

sig_df=df%>%filter(`FDR q-val` <=p)
  
  sig_df=sig_df%>%dplyr::select(NAME, NES, `FDR q-val`, gene_ratio)
  colnames(sig_df)=paste0(colnames(sig_df), '_GSEA')
  sig_df=sig_df%>%dplyr::rename('pathway'='NAME_GSEA')
  return(sig_df)
}
```

```{r fishers Method}
fishersMethod = function(x) pchisq(-2 * sum(log(x)),df=2*length(x),lower=FALSE)
```

```{r}
fisher_combine=function(df){
  ps=df %>% dplyr::select(padj, `FDR q-val_GSEA` )
  ps=ps %>% mutate(padj_avg=apply(ps, MARGIN = 1, fishersMethod))
  
  ps=ps%>% mutate(padj_avg=case_when(
  is.na(padj_avg)== TRUE~ padj, 
  .default = padj_avg
))
  
  ps=ps%>% mutate(padj_avg=case_when(
  is.na(padj_avg)== TRUE~ `FDR q-val_GSEA`,
  .default = padj_avg
))
 
  df=df%>%mutate(padj_avg=ps$padj_avg)
  
  return(df)
}
```

```{r}
mega_pathway_analysis=function(df, group, ont=c('GO', 'KEGG'), p=0.05, path_up, path_down){

############ FGSEA RESULTS
if(ont=='GO'){
  fgsea_up=as.data.frame(counts_to_fgsea_GO(df, p = p)[1])
  fgsea_down=as.data.frame(counts_to_fgsea_GO(df, p = p)[2])
} else if(ont=='KEGG'){
  fgsea_up=as.data.frame(counts_to_fgsea_KEGG(df, p = p)[1])
  fgsea_down=as.data.frame(counts_to_fgsea_KEGG(df, p = p)[2])
}

############ GSEA DESKTOP RESULTS
gsea_up=read_tsv(path_up)
gsea_down=read_tsv(path_down)
gsea_up=format_gsea_results(gsea_up, p=p)
gsea_down=format_gsea_results(gsea_down, p=p)

############ FORMAT KEGG ID FOR MERGE
if(ont=='KEGG'){
  gsea_up$pathway=tolower(gsea_up$pathway)
  gsea_down$pathway=tolower(gsea_down$pathway)
}

############ PREVENT ZERO RESULTS FROM CRASHING MERGE
if(dim(fgsea_up)[1]==0){
  fgsea_up[1, ]=NA
}else if(dim(fgsea_down)[1]==0){
  fgsea_down[1, ]=NA
}else if(dim(gsea_up)[1]==0){
  gsea_up[1, ]=NA
}else if(dim(gsea_down)[1]==0){
  gsea_down[1, ]=NA
}

full_down=full_join(fgsea_down, gsea_down)
full_up=full_join(fgsea_up, gsea_up)

############ FISHERS METHOD -- COMBINE P VALUES
full_down=fisher_combine(full_down)
full_up=fisher_combine(full_up)

############ COMBINE OTHER VARIABLES

full_down=full_down%>%
  #mutate(padj_avg=fishersMethod(across(c(padj, `FDR q-val_GSEA`))))%>%
  #mutate(padj_avg = rowMeans(across(c(padj, `FDR q-val_GSEA`)), na.rm=TRUE)) %>%
  mutate(NES_avg = rowMeans(across(c(NES, NES_GSEA)), na.rm=TRUE)) %>% 
  mutate(gene_ratio_avg=rowMeans(across(c(gene_ratio, gene_ratio_GSEA)), na.rm=TRUE))

full_up=full_up%>%
  #mutate(padj_avg=fishersMethod(across(c(padj, `FDR q-val_GSEA`))))
  #mutate(padj_avg = rowMeans(across(c(padj, `FDR q-val_GSEA`)), na.rm=TRUE)) %>%
  mutate(NES_avg = rowMeans(across(c(NES, NES_GSEA)), na.rm=TRUE)) %>%
  mutate(gene_ratio_avg=rowMeans(across(c(gene_ratio, gene_ratio_GSEA)), na.rm=TRUE))

########### EXPORT DATAFRAMES
# full_down=full_down%>%relocate(padj_avg, .after = pathway)%>%relocate(NES_avg, .after = padj_avg)
# full_up=full_up%>%relocate(padj_avg, .after = pathway)%>%relocate(NES_avg, .after = padj_avg)

full_down=full_down%>%relocate(NES_avg, .after = pathway)%>%relocate(padj_avg, .after = NES_avg)
full_up=full_up%>%relocate(NES_avg, .after = pathway)%>%relocate(padj_avg, .after = NES_avg)

filename=paste0('GSEA/', group, '/', group, '_', ont)

write.table(full_down, file=paste0(filename, '_full_down.txt'), quote=FALSE, row.names = FALSE)
write.table(full_up, file=paste0(filename, '_full_up.txt'), quote=FALSE, row.names = FALSE)

return(list(full_up, full_down))
}
```

```{r}
go_trim_merger=function(df_trim_up, df_trim_down, df_full_up, df_full_down){
  df_trim_up=df_trim_up%>% dplyr::rename('pathway'='TermID')
  df_trim_down=df_trim_down%>% dplyr::rename('pathway'='TermID')
  
  complete_up=inner_join(df_full_up, df_trim_up)
  complete_down=inner_join(df_full_down, df_trim_down)
  
  complete_up=complete_up%>% 
    mutate(padj_avg=case_when(padj_avg==0 ~.000000001, .default = padj_avg)) %>% 
    mutate(p_transform=-log(padj_avg, base = 10))%>%
    relocate(p_transform, .after = padj_avg)
  
  complete_down=complete_down%>% 
    mutate(padj_avg=case_when(padj_avg==0 ~.000000001, .default = padj_avg)) %>%
    mutate(p_transform=-log(padj_avg, base = 10))%>%
    relocate(p_transform, .after = padj_avg)

  return(list(complete_up, complete_down))
}
```

```{r}
kegg_formatter=function(df){
  df=df%>% 
    mutate(padj_avg=case_when(padj_avg==0 ~.000000001, .default = padj_avg)) %>% 
    mutate(p_transform=-log(padj_avg, base = 10))%>%
    relocate(p_transform, .after = padj_avg)

up_names=pathways.list[df$pathway]
pathway_id=names(up_names)
pathway_name=unname(up_names)
pathway_name=gsub(' - Staphylococcus epidermidis ATCC 12228', '', pathway_name)
pathway=data.frame("pathway"=pathway_id,
                      'pathway_name'=pathway_name)
df=inner_join(pathway, df, by='pathway')

df=df%>%dplyr::rename('Name'='pathway_name')

return(df)
}
```

```{r}
pathway_visualizer=function(df, group, ont=c('GO', 'KEGG'), direction=c('up', 'down')){
  
  df <- df[!is.na(df$pathway), ] 
  
  if(direction=='down'){
    title=paste(group, 'Significantly Suppressed Pathways' )
  } else if(direction=='up'){
    title=paste(group, 'Significantly Enriched Pathways' )
  }

  sub=paste(ont, 'database')
  
    p=ggplot(df, aes(x=p_transform, y=Name)) + 
    geom_point(aes(colour=gene_ratio_avg), size=3) + 
    geom_point(shape = 1,size = 3,colour = "black") +
  labs(title=title, subtitle = sub) +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
        theme(plot.subtitle = element_text(face='italic')) +
   theme_bw() + xlab('p value (-log10 scale)') + ylab('Significant Pathways') +
 #geom_text(aes(label = percent, vjust = 0.5, hjust=-0.5)) + 
  scale_size(guide="none") +
  scale_color_continuous(low='lightgray', high='darkgreen')
    
  plotname=paste(group, ont, direction, sep = '_')
  png(filename=paste0('GSEA/figures/', plotname, '.png'), units = 'in', height = 9, width = 9, res = 300)
  print(p)
  dev.off()
}
```

```{r}
plot_combiner=function(df_go, df_kegg, group, direction=c('up', 'down')){
  
  df_go <- df_go[!is.na(df_go$pathway), ] 
  df_kegg <- df_kegg[!is.na(df_kegg$pathway), ] 
  
  if(direction=='down'){
    title=paste(group, 'Significantly Suppressed Pathways' )
  } else if(direction=='up'){
    title=paste(group, 'Significantly Enriched Pathways' )
  }

  go=df_go %>% dplyr::select(pathway,Name, padj_avg, p_transform, NES_avg, gene_ratio_avg)
  go=go%>%mutate('db'='GO')

  kegg=df_kegg%>%dplyr::select(pathway, Name, padj_avg,p_transform, NES_avg, gene_ratio_avg )
  kegg=kegg%>%mutate('db'='KEGG')

  combined=rbind(go, kegg)
  
  if(direction=='up'){
  combined=combined %>%arrange(p_transform) %>% mutate(Name=factor(Name, levels=Name)) 
  } else if (direction=='down'){
  combined=combined %>%arrange(desc(p_transform)) %>% mutate(Name=factor(Name, levels=Name)) 
  }

  p=ggplot(combined, aes(x=p_transform, y=Name)) + 
  geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) + 
  geom_point(shape = as.factor(combined$db),size = 3,colour = "black") +
  labs(title=title) +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
        
  theme_bw() + xlab('p value (-log10 scale)') + ylab('Significant Pathways') +
 #geom_text(aes(label = percent, vjust = 0.5, hjust=-0.5)) + 
  scale_size(guide="none") +
  scale_color_continuous(low='lightgray', high='darkgreen', name = "Gene Ratio")+
        guides(shape = guide_legend(title = "Database"))
  
  print(p)
  
  # plotname=paste(group, 'both', direction, sep = '_')
  # png(filename=paste0('GSEA/figures/', plotname, '.png'), units = 'in', height = 9, width = 9, res = 300)
  # print(p)
  # dev.off()
  
  return(combined)
}
```

# GSEA desktop software
```
avg_counts=read.table('normalized_counts.txt')

avg_counts$gene=rownames(avg_counts)

avg_counts$gene=gsub('_C', '', avg_counts$gene)
avg_counts$gene=gsub('_P', '', avg_counts$gene)
avg_counts$gene=gsub('lepBP', 'lepB', avg_counts$gene)
avg_counts=avg_counts%>%dplyr::rename('gene_name'='gene')

count_table=inner_join(full_map, avg_counts)

count_table=count_table%>%dplyr::select(-c(gene_id, qseqid, length, pident, GO_terms, KEGG, 'gene name'))

count_table=count_table %>% mutate_if(is.numeric,
            round,
            digits = 0) %>%
  mutate_if(is.numeric,
            as.integer) %>%rename('Description'='gene_name')

count_table_kegg=count_table %>% dplyr::select(-uniprot_quickGO) %>% relocate(sseqid)
count_table_uniprot=count_table %>% dplyr::select(-sseqid) %>% relocate(uniprot_quickGO)

#count_table_uniprot_unique=count_table_uniprot%>%distinct(uniprot_quickGO, .keep_all = TRUE)
#count_table_kegg_unique=count_table_kegg%>%distinct(sseqid, .keep_all = TRUE)

write_delim(count_table_uniprot, 'GSEA/count_table_uniprot.txt', delim = "\t")
write_delim(count_table_kegg, 'GSEA/count_table_kegg.txt', delim = "\t")
```

```
avg_counts=read.table('heatshock_normalized_counts.txt')

avg_counts$gene=rownames(avg_counts)

avg_counts$gene=gsub('_C', '', avg_counts$gene)
avg_counts$gene=gsub('_P', '', avg_counts$gene)
avg_counts$gene=gsub('lepBP', 'lepB', avg_counts$gene)
avg_counts=avg_counts%>%dplyr::rename('gene_name'='gene')

count_table=inner_join(full_map, avg_counts)

count_table=count_table%>%dplyr::select(-c(gene_id, qseqid, length, pident, GO_terms, KEGG, 'gene name'))

count_table=count_table %>% mutate_if(is.numeric,
            round,
            digits = 0) %>%
  mutate_if(is.numeric,
            as.integer) %>%dplyr::rename('Description'='gene_name')

count_table_kegg=count_table %>% dplyr::select(-uniprot_quickGO) %>% relocate(sseqid)
count_table_uniprot=count_table %>% dplyr::select(-sseqid) %>% relocate(uniprot_quickGO)

#count_table_uniprot_unique=count_table_uniprot%>%distinct(uniprot_quickGO, .keep_all = TRUE)
#count_table_kegg_unique=count_table_kegg%>%distinct(sseqid, .keep_all = TRUE)

write_delim(count_table_uniprot, 'GSEA/heatshock_count_table_uniprot.txt', delim = "\t")
write_delim(count_table_kegg, 'GSEA/heatshock_count_table_kegg.txt', delim = "\t")
```

# G2 Analysis
### GO analysis
go_gmt has pathway name as a GO term, and the genes as uniprot
kegg_gmt has pathway name as a KEGG pathway, and genes as kegg genes

```{r}
pathup='GSEA/G2/G2_GO_gsea_up.tsv'
pathdown='GSEA/G2/G2_GO_gsea_down.tsv'

G2_go_full_up=as.data.frame(mega_pathway_analysis(G2_G0, group = 'G2', ont = 'GO', path_up = pathup, path_down = pathdown)[1])
G2_go_full_down=as.data.frame(mega_pathway_analysis(G2_G0, group = 'G2', ont = 'GO', path_up = pathup, path_down = pathdown)[2])
```

```{r revigo}
G2_trimmed_down=read_tsv('GSEA/G2/G2_trimmed_down.tsv')
G2_trimmed_up=read_tsv('GSEA/G2/G2_trimmed_up.tsv')

G2_go_up=as.data.frame(go_trim_merger(G2_trimmed_up, G2_trimmed_down, G2_go_full_up, G2_go_full_down)[1])
G2_go_down=as.data.frame(go_trim_merger(G2_trimmed_up, G2_trimmed_down, G2_go_full_up, G2_go_full_down)[2])
```

```{r visualizations}
pathway_visualizer(G2_go_down, group = 'G2', ont = 'GO', direction = 'down')
pathway_visualizer(G2_go_up, group = 'G2', ont = 'GO', direction = 'up')
```

### KEGG analysis
```{r}
pathup='GSEA/G2/G2_KEGG_gsea_up.tsv'
pathdown='GSEA/G2/G2_KEGG_gsea_down.tsv'

G2_kegg_full_up=as.data.frame(mega_pathway_analysis(G2_G0, group = 'G2', ont = 'KEGG', path_up = pathup, path_down = pathdown)[1])
G2_kegg_full_down=as.data.frame(mega_pathway_analysis(G2_G0, group = 'G2', ont = 'KEGG', path_up = pathup, path_down = pathdown)[2])

G2_kegg_up=kegg_formatter(G2_kegg_full_up)
G2_kegg_down=kegg_formatter(G2_kegg_full_down)

pathway_visualizer(G2_kegg_up, group = 'G2', ont = 'KEGG', direction = 'up')
pathway_visualizer(G2_kegg_down, group = 'G2', ont = 'KEGG', direction = 'down')
```


### Combined Plot

```{r}
G2_combined_down=plot_combiner(G2_go_down, G2_kegg_down, group = 'G2', direction = 'down')
G2_combined_up=plot_combiner(G2_go_up, G2_kegg_up, group = 'G2', direction = 'up')
```

```{r}
G2_combined_up[G2_combined_up$pathway=='sep00240', "db"] <- 'both'
G2_combined_up <- G2_combined_up[-2,]
```

```{r}
p=ggplot(G2_combined_up, aes(x=p_transform, y=Name)) + 
  geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) + 
  geom_point(shape = as.factor(G2_combined_up$db),size = 3,colour = "black") +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
        
  theme_bw() + xlab('p value (-log10 scale)') + ylab('Significant Pathways') +
 #geom_text(aes(label = percent, vjust = 0.5, hjust=-0.5)) + 
  scale_size(guide="none") +
  scale_color_continuous(low='lightgray', high='darkgreen', name = "Gene Ratio")+
        guides(shape = guide_legend(title = "Database"))
  
  print(p)
```

```
G2_combined_up %>%
  mutate(db = fct_relevel(db, 
            'GO', 'KEGG', 'both')) %>%
  ggplot( aes(x=p_transform, y=Name)) +
    geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) 
```

```
p=ggplot(G2_combined_up, aes(x=p_transform, y=Name)) + 
  geom_point(color='black', fill=G2_combined_up$gene_ratio_avg, size=3, aes(shape=factor(G2_combined_up$db))) + 
  scale_shape_manual(values=c(22, 21)) +
  scale_color_continuous(low='lightgray', high='darkgreen', name = "Gene Ratio")+

 #  #geom_point(shape = as.factor(G2_combined_up$db),size = 3,colour = "black") +


  #labs(title=title) +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
        
  theme_bw() + xlab('p value (-log10 scale)') + ylab('Significant Pathways') +
 #geom_text(aes(label = percent, vjust = 0.5, hjust=-0.5)) + 
  scale_size(guide="none") +
 # scale_color_continuous(low='lightgray', high='darkgreen', name = "Gene Ratio")+
        guides(shape = guide_legend(title = "Database"))
  
  print(p)

 #geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) + 
  # 
  # 
  # geom_point(color='black', shape=21, size=4, aes(fill=factor(group))) + 
  # scale_fill_manual(values=c('pink', 'lightgreen'))
```

spacer


spacer








# G5 Analysis
### GO analysis
go_gmt has pathway name as a GO term, and the genes as uniprot
kegg_gmt has pathway name as a KEGG pathway, and genes as kegg genes

```{r}
G5_G0=read_csv('G5_G0.csv')
pathup='GSEA/G5/G5_GO_gsea_up.tsv'
pathdown='GSEA/G5/G5_GO_gsea_down.tsv'


G5_go_full_up=as.data.frame(mega_pathway_analysis(G5_G0, group = 'G5', ont = 'GO', path_up = pathup, path_down = pathdown)[1])
G5_go_full_down=as.data.frame(mega_pathway_analysis(G5_G0, group = 'G5', ont = 'GO', path_up = pathup, path_down = pathdown)[2])
```

```{r revigo}
G5_trimmed_down=read_tsv('GSEA/G5/G5_trimmed_down.tsv')
G5_trimmed_up=read_tsv('GSEA/G5/G5_trimmed_up.tsv')

G5_go_up=as.data.frame(go_trim_merger(G5_trimmed_up, G5_trimmed_down, G5_go_full_up, G5_go_full_down)[1])
G5_go_down=as.data.frame(go_trim_merger(G5_trimmed_up, G5_trimmed_down, G5_go_full_up, G5_go_full_down)[2])
```

```{r visualizations}
pathway_visualizer(G5_go_down, group = 'G5', ont = 'GO', direction = 'down')
pathway_visualizer(G5_go_up, group = 'G5', ont = 'GO', direction = 'up')
```

### KEGG analysis
```{r}
pathup='GSEA/G5/G5_KEGG_gsea_up.tsv'
pathdown='GSEA/G5/G5_KEGG_gsea_down.tsv'

G5_kegg_full_up=as.data.frame(mega_pathway_analysis(G5_G0, group = 'G5', ont = 'KEGG', path_up = pathup, path_down = pathdown)[1])
G5_kegg_full_down=as.data.frame(mega_pathway_analysis(G5_G0, group = 'G5', ont = 'KEGG', path_up = pathup, path_down = pathdown)[2])

G5_kegg_up=kegg_formatter(G5_kegg_full_up)
G5_kegg_down=kegg_formatter(G5_kegg_full_down)

pathway_visualizer(G5_kegg_up, group = 'G5', ont = 'KEGG', direction = 'up')
pathway_visualizer(G5_kegg_down, group = 'G5', ont = 'KEGG', direction = 'down')
```


### Combined Plot

```{r}
G5_combined_down=plot_combiner(G5_go_down, G5_kegg_down, group = 'G5', direction = 'down')
G5_combined_up=plot_combiner(G5_go_up, G5_kegg_up, group = 'G5', direction = 'up')
```

spacer


spacer


# G10 Analysis
### GO analysis
go_gmt has pathway name as a GO term, and the genes as uniprot
kegg_gmt has pathway name as a KEGG pathway, and genes as kegg genes

```{r}
pathup='GSEA/G10/G10_GO_gsea_up.tsv'
pathdown='GSEA/G10/G10_GO_gsea_down.tsv'

G10_go_full_up=as.data.frame(mega_pathway_analysis(G10_G0, group = 'G10', ont = 'GO', path_up = pathup, path_down = pathdown)[1])
G10_go_full_down=as.data.frame(mega_pathway_analysis(G10_G0, group = 'G10', ont = 'GO', path_up = pathup, path_down = pathdown)[2])
```

```{r revigo}
G10_trimmed_down=read_tsv('GSEA/G10/G10_trimmed_down.tsv')
G10_trimmed_up=read_tsv('GSEA/G10/G10_trimmed_up.tsv')

G10_go_up=as.data.frame(go_trim_merger(G10_trimmed_up, G10_trimmed_down, G10_go_full_up, G10_go_full_down)[1])
G10_go_down=as.data.frame(go_trim_merger(G10_trimmed_up, G10_trimmed_down, G10_go_full_up, G10_go_full_down)[2])

pathway_visualizer(G10_go_down, group = 'G10', ont = 'GO', direction = 'down')
pathway_visualizer(G10_go_up, group = 'G10', ont = 'GO', direction = 'up')
```

### KEGG analysis
```{r}
pathup='GSEA/G10/G10_KEGG_gsea_up.tsv'
pathdown='GSEA/G10/G10_KEGG_gsea_down.tsv'

G10_kegg_full_up=as.data.frame(mega_pathway_analysis(G10_G0, group = 'G10', ont = 'KEGG', path_up = pathup, path_down = pathdown)[1])
G10_kegg_full_down=as.data.frame(mega_pathway_analysis(G10_G0, group = 'G10', ont = 'KEGG', path_up = pathup, path_down = pathdown)[2])

G10_kegg_up=kegg_formatter(G10_kegg_full_up)
G10_kegg_down=kegg_formatter(G10_kegg_full_down)

pathway_visualizer(G10_kegg_up, group = 'G10', ont = 'KEGG', direction = 'up')
pathway_visualizer(G10_kegg_down, group = 'G10', ont = 'KEGG', direction = 'down')
```

### Combined Plot

```{r}
plot_combiner(G10_go_down, G10_kegg_down, group = 'G10', direction = 'down')
G10_combined_up=plot_combiner(G10_go_up, G10_kegg_up, group = 'G10', direction = 'up')
```

# G10 beautification

```{r}
G10_combined_up=G10_combined_up[G10_combined_up$pathway!='GO:0055086', ]
G10_combined_up=G10_combined_up[G10_combined_up$pathway!='GO:0043412', ]
G10_combined_up$Name=as.character(G10_combined_up$Name)
```

```{r}
G10_combined_up$Name[G10_combined_up$Name=='RNA processing']='primary RNA processing'
G10_combined_up$Name[G10_combined_up$Name=='transposition']='transposition of DNA segments'
G10_combined_up
```
```{r}
G10_combined_up$Name=str_to_title(G10_combined_up$Name)
```
```{r}
G10_combined_up
```
```{r}
G10_combined_up$Name[G10_combined_up$Name=='Transposition Of Dna Segments']='Transposition of DNA Segments'
G10_combined_up$Name[G10_combined_up$Name=='Rna Modification']='RNA Modification'
G10_combined_up$Name[G10_combined_up$Name=='Primary Rna Processing']='Primary RNA Processing'
```


```{r}
G10_combined_up=G10_combined_up %>%arrange(p_transform) %>% mutate(Name=factor(Name, levels=Name)) 
```

```{r}
png('GSEA/pretty_G10_up.png', units = 'in', height = 2.75, width = 8, res = 300)
 p=ggplot(G10_combined_up, aes(x=p_transform, y=Name)) + 
  geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) + 
  geom_point(shape = as.factor(G10_combined_up$db),size = 3,colour = "black") +
  labs(title='Significantly Enriched Pathways', subtitle = '10 Mm Glucose') +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
        
  theme_bw() + xlab('p value (-log10 scale)') + 
   ylab('') +
 #geom_text(aes(label = percent, vjust = 0.5, hjust=-0.5)) + 
  scale_size(guide="none") +
  #scale_color_continuous(low='lightgray', high='darkgreen', name = "Gene Ratio")+
   #scale_color_gradient2(low='gold', mid = 'lightgreen', high='darkgreen', midpoint = 0.6, name = "Gene Ratio")+
   # scale_color_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12")) + 
           guides(shape = guide_legend(title = "Database", order = 2))+
    scale_color_distiller(palette = "YlOrBr", direction = 1, name='Gene Ratio')

 print(p)
 dev.off()
```


spacer

spacer



# G17 Analysis
### GO analysis
go_gmt has pathway name as a GO term, and the genes as uniprot
kegg_gmt has pathway name as a KEGG pathway, and genes as kegg genes

```{r}
pathup='GSEA/G17/G17_GO_gsea_up.tsv'
pathdown='GSEA/G17/G17_GO_gsea_down.tsv'

G17_go_full_up=as.data.frame(mega_pathway_analysis(G17_G0, group = 'G17', ont = 'GO', path_up = pathup, path_down = pathdown)[1])
G17_go_full_down=as.data.frame(mega_pathway_analysis(G17_G0, group = 'G17', ont = 'GO', path_up = pathup, path_down = pathdown)[2])
```

```{r revigo}
G17_trimmed_down=read_tsv('GSEA/G17/G17_trimmed_down.tsv')
G17_trimmed_up=read_tsv('GSEA/G17/G17_trimmed_up.tsv')

G17_go_up=as.data.frame(go_trim_merger(G17_trimmed_up, G17_trimmed_down, G17_go_full_up, G17_go_full_down)[1])
G17_go_down=as.data.frame(go_trim_merger(G17_trimmed_up, G17_trimmed_down, G17_go_full_up, G17_go_full_down)[2])

pathway_visualizer(G17_go_down, group = 'G17', ont = 'GO', direction = 'down')
pathway_visualizer(G17_go_up, group = 'G17', ont = 'GO', direction = 'up')
```

### KEGG analysis
```{r}
pathup='GSEA/G17/G17_KEGG_gsea_up.tsv'
pathdown='GSEA/G17/G17_KEGG_gsea_down.tsv'

G17_kegg_full_up=as.data.frame(mega_pathway_analysis(G17_G0, group = 'G17', ont = 'KEGG', path_up = pathup, path_down = pathdown)[1])
G17_kegg_full_down=as.data.frame(mega_pathway_analysis(G17_G0, group = 'G17', ont = 'KEGG', path_up = pathup, path_down = pathdown)[2])

G17_kegg_up=kegg_formatter(G17_kegg_full_up)
G17_kegg_down=kegg_formatter(G17_kegg_full_down)

pathway_visualizer(G17_kegg_up, group = 'G17', ont = 'KEGG', direction = 'up')
pathway_visualizer(G17_kegg_down, group = 'G17', ont = 'KEGG', direction = 'down')
```
### Combined Plot

```{r}
G17_combined_down=plot_combiner(G17_go_down, G17_kegg_down, group = 'G17', direction = 'down')
G17_combined_up=plot_combiner(G17_go_up, G17_kegg_up, group = 'G17', direction = 'up')
```
```{r}
G17_combined_down
```


# G17 beautification

```{r}
G17_combined_down=G17_combined_down[G17_combined_down$pathway!='GO:0006796', ]

G17_combined_down=G17_combined_down[G17_combined_down$pathway!='GO:0044281', ]

G17_combined_down=G17_combined_down[G17_combined_down$pathway!='GO:0009056', ]

G17_combined_down=G17_combined_down[G17_combined_down$pathway!='GO:0044238', ]

G17_combined_down$Name=as.character(G17_combined_down$Name)
```

```{r}
G17_combined_down$Name[G17_combined_down$Name=='Two-component system']='Two-component signal transduction system'
G17_combined_down$Name[G17_combined_down$Name=='Ribosome']='Ribosome/tRNA translation'

G17_combined_down
```

```{r}
G17_combined_down$Name=str_to_title(G17_combined_down$Name)
```

```{r}
G17_combined_down$Name[G17_combined_down$Name=='Citrate Cycle (Tca Cycle)']='Citrate Cycle (TCA Cycle)'
G17_combined_down$Name[G17_combined_down$Name=='Ribosome/Trna Translation']='Ribosome / tRNA Translation'
G17_combined_down$Name[G17_combined_down$Name=='Energy Derivation By Oxidation Of Organic Compounds']='Energy Derivation by Oxidation of Organic Compounds'
G17_combined_down$Name[G17_combined_down$Name=='Glycine, Serine And Threonine Metabolism']='Glycine, Serine & Threonine Metabolism'
G17_combined_down$Name[G17_combined_down$Name=='Cysteine And Methionine Metabolism']='Cysteine & Methionine Metabolism'
```

```{r}
G17_combined_down
```

```{r}
G17_combined_down=G17_combined_down %>%arrange(desc(p_transform)) %>% mutate(Name=factor(Name, levels=Name)) 
```


```{r}
png('GSEA/pretty_G17_down.png', units = 'in', height = 7, width = 8, res = 300)

p= ggplot(G17_combined_down, aes(x=p_transform, y=Name)) + 
  geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) + 
  geom_point(shape = as.factor(G17_combined_down$db),size = 3,colour = "black") +
      guides(shape = guide_legend(title = "Database"))+
            scale_color_distiller(direction = 1, name='Gene Ratio')+  
  labs(title='Significantly Suppressed Pathways', subtitle = '17 Mm Glucose') +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
  theme_bw() + xlab('p value (-log10 scale)') + ylab('') +
  scale_size(guide="none") 
print(p)
dev.off()
```
# Drew Edits

1) If we can edit pathway names then I’d offer the following upgrades:

— anything that is “Metabolic Process” could be “Metabolism”; what do you think?
— anything that is “Biosynthetic Process” could be “Biosynthesis”; ditto?

```{r}
G17_combined_down$Name=gsub('Metabolic Process', 'Metabolism', G17_combined_down$Name)
G17_combined_down$Name=gsub('Biosynthetic Process', 'Biosynthesis', G17_combined_down$Name)

G10_combined_up$Name=gsub('Metabolic Process', 'Metabolism', G10_combined_up$Name)
G10_combined_up$Name=gsub('Biosynthetic Process', 'Biosynthesis', G10_combined_up$Name)
```
(2) do we know what “Gene Expression” refer to?  E.g., is this genes encoding RNA polymerase?  Sigma factors?  Other?


(3) “Gene Ratio” for fraction of pathway genes differentially regulated is a potentially confusing label.  I.e., b/c we are referring to fold change throughout (w/r/t/ differential gene expression) and that is also a ratio I worry that gene ratio here doesn’t clearly make the point.  “Percentage of Pathway Genes” and using 0-100% may work better; what do you think?
```{r}
G17_combined_down=G17_combined_down%>%mutate(gene_ratio_avg=round(gene_ratio_avg*100))
G17_combined_down=G17_combined_down %>%arrange(desc(p_transform)) %>% mutate(Name=factor(Name, levels=Name)) 


G10_combined_up=G10_combined_up%>%mutate(gene_ratio_avg=round(gene_ratio_avg*100))
G10_combined_up=G10_combined_up %>%arrange(p_transform) %>% mutate(Name=factor(Name, levels=Name)) 
```

```{r}
G17_combined_down
```

```{r}
# Checking Gene Expression Genes
  SD_formatted=deseq_format(G17_G0)
  ranks=SD_formatted$log2FoldChange
  names(ranks)=SD_formatted$uniprot_quickGO
  ranks=sort(ranks, decreasing = TRUE)
```
  
```{r}
  set.seed(12345)
  fgsea_res=fgsea(go_gmt, ranks, minSize = 15, maxSize = 500)
```

```{r}
GE_genes_SD=unlist(fgsea_res[fgsea_res$pathway=='GO:0010467', ]$leadingEdge)
```

```{r}
GE_genes_SD
```
```{r}
GE_genes_SD=full_map[full_map$uniprot_quickGO %in%GE_genes_SD, ]
GE_genes_SD=GE_genes_SD%>%dplyr::select(gene_name, gene_id)

GE_genes_G17=inner_join(GE_genes_SD, SD_formatted)%>%arrange(desc(log2FoldChange))%>%dplyr::select(gene_name, gene_id)
```

```{r}
write.csv(GE_genes_G17, 'G17_gene_expression_genes.csv')
```


(4) I *really* like your aesthetic choices on the plot for repressed genes; let’s keep this; can we make the spacing exactly the same for the upregulated genes.

```{r}
G17_combined_down[1:5, 4]=5

```

```{r}
G10_combined_up[5, 4]=5
```


```{r}
#png('GSEA/pretty_G17_down.png', units = 'in', height = 7, width = 8, res = 300)
G17_combined_down=G17_combined_down %>%arrange(desc(p_transform)) %>% mutate(Name=factor(Name, levels=Name)) 
p= ggplot(G17_combined_down, aes(x=p_transform, y=Name)) + 
  geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) + 
  geom_point(shape = as.factor(G17_combined_down$db),size = 3,colour = "black") +
      guides(shape = guide_legend(title = "Database"))+
            scale_color_distiller(direction = 1, name='Percentage of\nPathway Genes')+  
  labs(title='Significantly Suppressed Pathways', subtitle = '17 Mm Glucose') +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
  theme_bw() + xlab('p value (-log10 scale)') + ylab('') +
  scale_size(guide="none") +xlim(1,9)
print(p)
#dev.off()
```

```{r}
png('GSEA/pretty_G10_up.png', units = 'in', height = 2.75, width = 8, res = 300)
 p=ggplot(G10_combined_up, aes(x=p_transform, y=Name)) + 
  geom_point(aes(colour=gene_ratio_avg, shape=db), size=3) + 
  geom_point(shape = as.factor(G10_combined_up$db),size = 3,colour = "black") +
  labs(title='Significantly Enriched Pathways', subtitle = '10 Mm Glucose') +
  theme(axis.title = element_text(face="bold", size=20))+
  theme(axis.text.x= element_text(face='bold', size=20))+
  theme(axis.text.y= element_text(face='bold', size=20))+
        
  theme_bw() + xlab('p value (-log10 scale)') + 
   ylab('') +
 #geom_text(aes(label = percent, vjust = 0.5, hjust=-0.5)) + 
  scale_size(guide="none") +
  #scale_color_continuous(low='lightgray', high='darkgreen', name = "Gene Ratio")+
   #scale_color_gradient2(low='gold', mid = 'lightgreen', high='darkgreen', midpoint = 0.6, name = "Gene Ratio")+
   # scale_color_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12")) + 
           guides(shape = guide_legend(title = "Database", order = 2))+
  xlim(1,9)+
    scale_color_distiller(palette = "YlOrBr", direction = 1, name='Percentage of\nPathway Genes')

 print(p)
 dev.off()
```





# G50 Analysis 
### GO analysis
go_gmt has pathway name as a GO term, and the genes as uniprot
kegg_gmt has pathway name as a KEGG pathway, and genes as kegg genes

```{r}
G50_G0=read_csv('G50_G0.csv')
pathup='GSEA/G50/G50_GO_gsea_up.tsv'
pathdown='GSEA/G50/G50_GO_gsea_down.tsv'

G50_go_full_up=as.data.frame(mega_pathway_analysis(G50_G0, group = 'G50', ont = 'GO', path_up = pathup, path_down = pathdown)[1])
G50_go_full_down=as.data.frame(mega_pathway_analysis(G50_G0, group = 'G50', ont = 'GO', path_up = pathup, path_down = pathdown)[2])
```

```{r revigo}
G50_trimmed_down=read_tsv('GSEA/G50/G50_trimmed_down.tsv')
G50_trimmed_up=read_tsv('GSEA/G50/G50_trimmed_up.tsv')

G50_go_up=as.data.frame(go_trim_merger(G50_trimmed_up, G50_trimmed_down, G50_go_full_up, G50_go_full_down)[1])
G50_go_down=as.data.frame(go_trim_merger(G50_trimmed_up, G50_trimmed_down, G50_go_full_up, G50_go_full_down)[2])

pathway_visualizer(G50_go_down, group = 'G50', ont = 'GO', direction = 'down')
pathway_visualizer(G50_go_up, group = 'G50', ont = 'GO', direction = 'up')
```

### KEGG analysis
```{r}
pathup='GSEA/G50/G50_KEGG_gsea_up.tsv'
pathdown='GSEA/G50/G50_KEGG_gsea_down.tsv'

G50_kegg_full_up=as.data.frame(mega_pathway_analysis(G50_G0, group = 'G50', ont = 'KEGG', path_up = pathup, path_down = pathdown)[1])
G50_kegg_full_down=as.data.frame(mega_pathway_analysis(G50_G0, group = 'G50', ont = 'KEGG', path_up = pathup, path_down = pathdown)[2])

G50_kegg_up=kegg_formatter(G50_kegg_full_up)
G50_kegg_down=kegg_formatter(G50_kegg_full_down)

pathway_visualizer(G50_kegg_up, group = 'G50', ont = 'KEGG', direction = 'up')
pathway_visualizer(G50_kegg_down, group = 'G50', ont = 'KEGG', direction = 'down')
```

### Combined Plot

```{r}
plot_combiner(G50_go_down, G50_kegg_down, group = 'G50', direction = 'down')
plot_combiner(G50_go_up, G50_kegg_up, group = 'G50', direction = 'up')
```



spacer


spacer


# End
spacer

